---
title: "Oil Spill: Spatial"
description: |
  Exploratory interactive maps showing the location of oil spill events from 2008 data from CA DFW Oil Spill Incident Tracking.
author:
  - name: Elmera Azadpour
    url: {}
date: 2021-02-23
output:
  distill::distill_article:
    self_contained: false
---

## Load packages
```{r setup}
library(tidyverse)
library(here)
library(janitor)
library(sf)
library(tmap) # for interactive graph
library(gstat) # for kriging
library(stars) #grid for new location predictions
```

## Read in data
```{r, warning=FALSE}
ca_counties <- read_sf(here("_posts", "2021-02-23-oil-spill-spatial", "ca_counties"), layer = "CA_Counties_TIGER2016") %>% 
  clean_names() %>% 
  select(name)

st_crs(ca_counties) # "EPSG",3857


oil_spill <- read_sf(here("_posts", "2021-02-23-oil-spill-spatial", "oilspill"), layer = "Oil_Spill_Incident_Tracking_%5Bds394%5D") %>% 
  clean_names()

st_crs(oil_spill)  # "EPSG",3857

```

## exploratory interactive map in tmap showing the location of oil spill events included in the data
```{r, warning=FALSE}
ggplot() +
  geom_sf(data = ca_counties) +
  geom_sf(data = oil_spill, size = .5)
```

## Interactive map with tmap! 
```{r}
tmap_mode("view")

tm_shape(oil_spill) + 
  tm_dots()
```


## Static chloropleth map in ggplot in which the fill color for each county depends on the count of inland oil spill events by county for the 2008 oil spill data
```{r, warning=FALSE}
# oil_spill_inland <- oil_spill %>% 
#   filter(inlandmari %in% c("Inland")) %>% 
#   group_by(localecoun) %>% 
#   summarise(count = n())

county_oil <- ca_counties %>% 
  st_join(oil_spill) %>% 
   filter(inlandmari %in% c("Inland")) %>% 
  group_by(localecoun) %>% 
  summarise(count = n())
  
ggplot() +
  geom_sf(data = ca_counties) +
  geom_sf(data = county_oil, aes(fill = count)) +
  theme_bw()+ 
  scale_fill_gradient(low = "grey", high = "red") 
```

## Citation:
+ Oil spill data provided by: CA State Geoportal: Oil Sill Incident Tracking [ds394] https://gis.data.ca.gov/datasets/7464e3d6f4924b50ad06e5a553d71086_0/data