---
title: "Text wrangling and analysis"
description: |
   Counts and visualizations for the most frequently used words in "The Origin of Species" by Charles Darwin, 6th Edition.
author:
  - name: Elmera Azadpour
    url: {}
date: 2021-02-23
output:
  distill::distill_article:
    self_contained: false
---
## Load packages
```{r setup}
library(tidyverse)
library(here)
library(tidytext)
library(textdata)
library(pdftools)
library(ggwordcloud)
```

### To access data, html and Rmd files: 
+ https://github.com/elmeraa/244-Assignment3 
+ you will find the Origin of Species pdf in the "data" folder
+ you will find the Rmd and html files as "a3-task3" in the "src" folder

## Read in the Origin of Species, 6th Edition by Darwin 
```{r, warning=FALSE}
os_text <- pdf_text(here("_posts", "2021-02-23-text-wrangling-and-analysis", "originofspecies6th-darwin.pdf"))
```

## Text into a data frame, then  wrangling with the tidyverse, break it up by chapter, and do some analyses.
```{r, warning=FALSE}
os_tidy <- data.frame(os_text) %>% 
  mutate(text_full = str_split(os_text, pattern = '\\n')) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full))

```

```{r, warning=FALSE}
os_df <- os_tidy %>% 
  slice(-(1:184)) %>% 
mutate(chapter = case_when(
     str_detect(text_full, "CHAPTER") ~ str_extract(text_full, "CHAPTER [1-9]+"),
     TRUE ~ NA_character_
   )) %>% 
  fill(chapter) %>% 
  separate(col = chapter, into = c("cha", "no"), sep = " ") %>% 
  mutate(chapter = as.numeric(no))
```

## Word count by Chapter
```{r, warning=FALSE}
os_tokens <- os_df %>% 
  unnest_tokens(word, text_full) %>% 
  select(-os_text) 

os_tokens_clean <- os_tokens %>%
   mutate(word = str_replace(word, "[0-9-]+", NA_character_)) %>% 
  drop_na()

os_wordcount <- os_tokens_clean %>% 
  count(chapter, word) 
```


## Remove stop words and recounting again
```{r, warning=FALSE}
os_nonstop_words <- os_tokens_clean %>% 
  anti_join(stop_words)

nonstop_counts <- os_nonstop_words %>% 
  count(chapter, word) 
  
```

## Top 5 words by chapter
```{r, warning=FALSE}
top_5_words <- nonstop_counts %>% 
  group_by(chapter) %>% 
  arrange(-n) %>% 
  slice(1:5)

## vizualization
ggplot( data = top_5_words,
        aes(word,n )) +
  geom_col(fill = "blue") +
  facet_wrap(~chapter, scales = "free") +
  coord_flip()
```

## Word cloud for all text
```{r, warning=FALSE}
nonstop_counts_full <- os_nonstop_words %>% 
  count(word) 

os_top100_removesps <- nonstop_counts_full %>% 
  arrange(-n) %>% 
  slice(1:100)

os_cloud <- ggplot(data = os_top100_removesps, aes(label = word)) +
  geom_text_wordcloud(aes(color = n, size = n), shape = "circle") +
  scale_size_area(max_size = 10) +
  scale_color_gradient(low = "darkseagreen", high = "forestgreen") +
  theme_minimal()

os_cloud

#ggsave(here("src","originofspecies-wc-ea.png"), width = 8, height = 5) # to save
```

## Sentiment analysis using "NRC" lexicon
```{r, warning=FALSE}
## nrc to just check out
os_nrc <- os_nonstop_words %>% 
  inner_join(get_sentiments("nrc"))

os_nrc_counts <- os_nrc %>% 
  count(chapter, sentiment)

ch_names <- list(
  "1" ="Chapter 1",
  "2" = "Chapter 2",
  "3" = "Chapter 3",
  "4" = "Chapter 4",
  "5" = "Chapter 5",
  "6" = "Chapter 6",
  "7" = "Chapter 7",
  "8" = "Chapter 8",
  "9" = "Chapter 9",
  "10" = "Chapter 10",
  "11" = "Chapter 11",
  "12" = "Chapter 12",
  "13" = "Chapter 13",
  "14" = "Chapter 14",
  "15" = "Chapter 15"
 )

ch_labeller <- function(variable,value){
  return(ch_names[value])
}

ggplot(data = os_nrc_counts, aes(x = sentiment, y = n)) +
  geom_col() +
  facet_wrap(~chapter, labeller=ch_labeller) +
  coord_flip() +
  theme_minimal() +
  labs(y = "Word Count", x = "NRC sentiment", title = "Sentiment analysis of the Origin of Species by Charles Darwin\nusing NRC from Saif Mohammad and Peter Turney")
  
```


## Citations:
+ NRC lexicon: Crowdsourcing a Word-Emotion Association Lexicon, Saif Mohammad and Peter Turney, Computational Intelligence, 29 (3), 436-465, 2013.
+ Origin of Species Text: https://laurieximenez.files.wordpress.com/2019/04/the-origin-of-species_charles-darwin.pdf
