---
title: "Survey responses of political affiliation and recognition as an environmentalist with chi-square analysis"
description: |
  This script analyzes the association between respondents’ political affiliation (coined “conservative” or “liberal” by self-identification) and if they consider themself an environmentalist, followed by a chi-square test for independence.
author:
  - name: Elmera Azadpour
    url: {}
date: 2021-02-23
output:
  distill::distill_article:
    self_contained: false
---
## Load packages
```{r setup}
library(tidyverse)
library(here)
library(janitor)
library(broom)
library(kableExtra)
```

## Introduction:
This task analyzes the association between respondents’ political affiliation (coined “conservative” or “liberal” by self-identification) and if they consider themself an environmentalist. The variables in the dataset (conservation_survey.csv) I have used include:

- env_id: response to survey statement "I consider myself an environmentalist," with outcomes 1 = Yes, 2 = No, 3 = Unsure
- pol_or: response to survey question "How do you self-identify politically?" Response values 1 - 3 are “conservative” (strongly to slightly), and values 5 - 7 are “liberal” (slightly to strongly).

```{r, error=FALSE, warning=FALSE}
## Read in survey data 
survey <- read_csv(here("_posts", "2021-02-23-chisquare-politcal", "conservation_survey.csv"))
```

```{r, error=FALSE, warning=FALSE}
## only keep variables: 'env_id' and 'pol_or'
survey_clean <- survey %>% 
  clean_names() %>% 
  select(env_id, pol_or) %>% 
  mutate(party = case_when(
    between(pol_or, 1,3) ~ "conservative",
    between(pol_or, 5,7) ~ "liberal",
    TRUE ~ NA_character_))

survey_environmentalist <- survey_clean %>%  
  mutate(environmentalist = case_when(
   env_id == 1 ~ "yes",
   env_id == 2 ~ "no",
   env_id == 3 ~ "unsure"))
```

```{r, error=FALSE, warning=FALSE}
## Finalized table that shows counts and proportions of "liberal" and "conservative" survey respondents who responded with "yes", "no" and "unsure" to the statement: "I consider myself an environmentalist"
survey_counts <- survey_environmentalist %>% 
  janitor::tabyl(party, environmentalist)

survey_proportions <- survey_counts %>% 
  adorn_percentages() %>% 
  janitor::adorn_pct_formatting(digits = 2) %>% 
  adorn_ns() %>% 
  drop_na()

survey_ct <- survey_proportions %>% 
  column_to_rownames(var = "party")

survey_ct %>%  
  kable(col.names = c("No",
                      "Unsure",
                      "Yes"),
    caption = "**Table 1**: Association between respondents’ political affiliation (“conservative” or “liberal” by self-identification) vs if they consider themself an environmentalist") %>% 
  kable_styling(full_width = FALSE) 
```

```{r, error=FALSE, warning=FALSE}
## Chi-square test
survey_ct <- survey_counts %>%   
  drop_na() %>% 
  column_to_rownames(var = "party")
  
survey_x2 <- chisq.test(survey_ct)
survey_x2

survey_tidy <- tidy(survey_x2)
```
## Chi-square results and summary
A chi-square test for independence compares two variables in a contingency table to see if they are related. In a more general sense, it tests to see whether distributions of categorical variables differ from each another. For this analysis, there is a significant association (i.e. non-independence) between political affiliation and and if a person considers themself an environmentalist ($\chi$^2^(`r survey_tidy$parameter`) = `r round(survey_tidy$statistic,2)`, *p*-value = `r format(survey_tidy$p.value, scientific = TRUE, digits = 3)`). 


#### Citations:
+ Andrews Forest LTER Site and C. Batavia. 2019. Influences on charitable giving for conservation: Online survey data of 1,331 respondents across the US, August 2017 ver 3. Environmental Data Initiative. https://doi.org/10.6073/pasta/926e6270e324a1322a900da14d38b96c

+ Metadata, including survey questions and possible responses, are here: https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-and.5444.3
