---
title: "Species Richness of Cetaceans along Coastal California"
description: |
  Raster of species richness of cetacean species off the coast of California utlizing a probability threshold.
author:
  - name: Elmera Azadpour
    url: {}
date: 2021-03-07
output:
  distill::distill_article:
    self_contained: false
---

### To access data, html and Rmd files: 
+ https://github.com/elmeraa/244-Assignment3 
+ you will find the cetacean data in the "data" folder
+ you will find the Rmd and html files as "a3-task2" in the "src" folder

### Load Packages
```{r setup}
library(tidyverse)
library(here)
library(janitor)
library(sf)
library(raster)
library(fasterize)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
```

### Read in data and stack files
```{r}
cetacean <-  list.files(here("_posts", "2021-03-07-cetacean-richness","ca_cetaceans"), full.names = TRUE)
cetacean_raster <- stack(cetacean)
cetacean_raster
plot(cetacean_raster)
```

### Apply probability threshold (of 0.6 or greater) to determine presence or non-presence
```{r}
aquamap_fun <- function(x, thresh = 0.6){
  y <- ifelse(x >= thresh, 1, 0)
  return(y)
}

aquamap <- calc(cetacean_raster, fun = aquamap_fun)
species_richness1 <- calc(aquamap, fun = sum, na.rm = TRUE)
plot(species_richness1)
```

### Converting to dataframe and plotting
```{r, fig.cap= "**Fig. 1** Species Richness of 16 Cetacean Species along the Coast of California. Presence of Cetacean species was determined with with a probability threshold of 0.6 or greater."}
aquamap_df <- raster::rasterToPoints(species_richness1) %>% 
  as.data.frame()

states110 <- ne_download(scale = 110, type = 'states', category = 'cultural', returnclass = "sf") %>% filter(name %in% c("California"))

ggplot() +
geom_raster(data = aquamap_df, aes(x=x,y=y, fill = layer)) +
  geom_sf(data = states110) +
  scale_fill_gradient2(low = "red", mid = "white", high = "blue") +
  labs(x = "Longitude", y = "Latitude", title = "Species Richness of Cetaceans along Coastal California", fill = "Number of Cetacean Species") +
  theme_classic()
```

### Citation:
+ Kaschner, K., Rius-Barile, J., Kesner-Reyes, K., Garilao, C., Kullander, S., Rees, T., & Froese, R. (2016). AquaMaps: Predicted range maps for aquatic species. www.aquamaps.org
