---
title: "Hierarchical Clustering"
description: |
  Hierarchical clustering by complete linkage to create a dendrogram showing multivariate clustering for water chemistry by site in Santa Barbara, CA.
author:
  - name: Elmera Azadpour
    url: {}
date: 2021-03-07
output:
  distill::distill_article:
    self_contained: false
---

### To access data, html and Rmd files: 
+ https://github.com/elmeraa/244-assignment4
+ you will find the stream chemistry data in the "data" folder
+ you will find the Rmd and html files as "a4-task1" in the "src" folder


## Load packages
```{r setup}
library(tidyverse)
library(here)
library(janitor)
# Packages for cluster analysis:
library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)
```

### Introduction: 
+ To perform hierarchical clustering by site, we'll begin by making a data frame that has a single summary row per site (e.g. based on means from all observations at that site), then we will calculate the euclidean distance before performing complete linkage agglomerative hierarchical clustering. 

### Read in data
```{r, results = "hide"}
stream_chem <- read_csv(here("_posts", "2021-03-07-hierarchical-clustering","sbc_lter_registered_stream_chemistry.csv")) %>% 
  clean_names() %>% 
  na_if(-999.0) %>% 
  group_by(site_code) %>% 
  summarise(across(nh4_u_m:spec_cond_u_spercm, mean, na.rm = TRUE)) %>% 
  drop_na()
```

### Scale the data 
```{r}
# scale the numeric variables
stream_chem_scaled <- stream_chem %>% 
  select(2:11) %>% 
  scale()

# make rownames the site code names from original data set
rownames(stream_chem_scaled) <- stream_chem$site_code

# Compute dissimilarity values (Euclidean distances):
stream_euc_distance <- dist(stream_chem_scaled, method = "euclidean")

# Check out the output:
# View(euc_distance)
```

### Perform hierarchical clustering by complete linkage with `stats::hclust()`
```{r}
# Hierarchical clustering (complete linkage)
stream_hc_complete <- hclust(stream_euc_distance, method = "complete" )

# Plot it (base plot):
plot(stream_hc_complete, cex = 0.6, hang = -1)
```

### Convert to class: dendogram
```{r}
stream_dend_complete <- as.dendrogram(stream_hc_complete)
```

### Plot using `ggdendrogram()`, a `ggplot` wrapper: 
```{r}
# ggplot 
ggdendrogram(stream_dend_complete, 
             rotate = TRUE)+
  labs(x = "Santa Barbara area watershed site code") +
  ylab(NULL)+ 
  theme_bw() + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 
```

### Data & Metadata Source:
+ SBC LTER: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000
Creators: Santa Barbara Coastal LTER, & Melack, John M
Citation: Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative. https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174. 